Variana Reloaded

### 2023-05-08 cv/cv.tex
Update CV

\noindent Hands-on technical team leader, engineer at heart, PhD, with a specialist expertise in artificial intelligence and data analysis. Has operated in public research and private R\&D in various fields (healthcare, electro-domestic, entertainment). Strong experience in the development of both open source and commercial software in Python, C, C++. {\em Keywords:} applied mathematics, statistics, machine learning, computer vision, computer graphics, medical image analysis, clinical studies.

@@

@@

\ \\
\textbf{STATISTICS:} Strong experience developing inferential statistical models for machine learning, computer vision, computer graphics, and performing data analysis for cognitive research, clinical studies, software validation, \ldots

@@

\bighead{HEAD OF DIDIMO AUTOMATION TEAM}{Didimo, Porto, Portugal}{2019--present}
\nopagebreak
\begin{itemize}
\item Leading development of generation pipeline of high-fidelity avatars from photos. {\em Keywords:} facial reconstruction, ``deep face'', animation retargetting, image feature extraction, image restoration.
\end{itemize}
\ \\

@@

\begin{tabular}{p{0.2\linewidth}p{0.8\linewidth}}
  \textbf{Languages:} & Native French, fluent English, basic Portuguese \& Italian.\\
  %\hline
  \ \\
  \textbf{IT skills:} &
  \tabitem Programming languages: Python, C, C++, Matlab, R.\\
  & \tabitem Long-term experience with scientific Python packages (numpy, scipy, pylab).\\
  & \tabitem CV/AI Python packages: skimage, PIL, opencv, sklearn, pytorch, tensorflow, keras.\\
  & \tabitem  C/Python integration via Cython.\\
  & \tabitem Version control software: git, svn, perforce.\\
  & \tabitem Agile workflow (mainly Scrum) using JIRA \& Confluence.\\
  & \tabitem Reporting: \LaTeX, MS~Office, OpenOffice, LibreOffice, Google~Docs.\\
  %\hline
  \ \\
  \textbf{Publications:} &
  \tabitem ORCID: \url{http://orcid.org/0000-0002-4821-6893}. \\
  & \tabitem 116 scholarly publications.\\
  & \tabitem Google scholar statistics: 7308 citations, h-index: 35, i10-index: 72 (October 2022).\\
  & \tabitem 5 published international patents. 2 ongoing applications.\\
  %\hline
  \ \\
  \textbf{Academic work:} & 
  \tabitem Post-graduate teaching in computer vision and medical image analysis: about 100~hours (France: Ecole Centrale Paris, University of Nice, INSERM/CNRS continuing education; Switzerland: EPFL Lausanne, University of Lausanne).\\
  & \tabitem Long-time scientific journal reviewing experience (IEEE Trans.~Medical Imaging, Medical Image Analysis, NeuroImage, IEEE Trans.~PAMI, IEEE Trans.~Signal Processing, Frontiers in Neuroscience, \ldots).\\
  & \tabitem Supervision of two PhD students and four MSc students as main supervisor, three PhD students and two MSc students as co-supervisor.\\
  %\hline
  \ \\
  \textbf{References:} & Available on request.
\end{tabular}

>>>>

\noindent Hands-on technical team leader, engineer at heart, PhD, with a specialist expertise in computer vision and artificial intelligence. Has operated in public research and private R\&D in various fields (healthcare, electro-domestic, entertainment). Strong experience in the development of both open source and commercial software in Python, C, C++. {\em Keywords:} applied mathematics, statistics, machine learning, computer vision, computer graphics, computer-aided diagnosis, smart appliances, digital avatars.

@@

\ \\
\textbf{COMPUTATIONAL STATISTICS:} 25 years experience developing inferential statistical models for machine learning, computer vision, computer graphics, \ldots

@@

@@

\bighead{HEAD OF ARTIFICIAL INTELLIGENCE}{Didimo, Porto, Portugal}{2019--present}
\nopagebreak
\begin{itemize}
\item Leading development of generation pipeline of high-fidelity avatars from photos. {\em Keywords:} facial reconstruction, deep face, animation retargeting, image feature extraction, image restoration.
\end{itemize}
\ \\

@@

\begin{tabular}{p{0.2\linewidth}p{0.8\linewidth}}
  \textbf{Languages:} & Native French, fluent English, basic Portuguese \& Italian.\\
  %\hline
  \ \\
  \textbf{IT skills:} &
  \tabitem Programming languages: Python, C, C++, Matlab, R.\\
  & \tabitem Long-term experience with scientific Python packages (numpy, scipy, pylab).\\
  & \tabitem CV/AI Python packages: skimage, PIL, opencv, sklearn, pytorch, tensorflow, keras.\\
  & \tabitem  C/Python integration via Cython.\\
  & \tabitem Version control software: git, svn, perforce.\\
  & \tabitem Agile workflow (mainly Scrum) using JIRA \& Confluence.\\
  & \tabitem Reporting: \LaTeX, MS~Office, OpenOffice, LibreOffice, Google~Docs.\\
  %\hline
  \ \\
  \textbf{Publications:} &
  \tabitem ORCID: \url{http://orcid.org/0000-0002-4821-6893}. \\
  & \tabitem 116 scholarly publications.\\
  & \tabitem Google scholar statistics: 7603 citations, h-index: 35, i10-index: 74 (May 2023).\\
  & \tabitem 5 published international patents. 2 ongoing applications.\\
  %\hline
  \ \\
  \textbf{Academic work:} & 
  \tabitem Post-graduate teaching in computer vision and medical image analysis: about 100~hours (France: Ecole Centrale Paris, University of Nice, INSERM/CNRS continuing education; Switzerland: EPFL Lausanne, University of Lausanne).\\
  & \tabitem Long-time scientific journal reviewing experience (IEEE Trans.~Medical Imaging, Medical Image Analysis, NeuroImage, IEEE Trans.~PAMI, IEEE Trans.~Signal Processing, Frontiers in Neuroscience, \ldots).\\
  & \tabitem Supervision of two PhD students and four MSc students as main supervisor, three PhD students and two MSc students as co-supervisor.\\
  %\hline
  \ \\
  \textbf{References:} & Available on request.
\end{tabular}

### 2023-05-08 papers/variana/notes.tex
Update CV

>>>>

\section{2023 thoughts}

Context: ill posedness (in the sense of Hadamard) of empirical risk minimization, including  maximum likelihood, MAP, etc. These methods are at the heart of statistical inference and machine learning, still they are based on ill-posed mathematical problems! Shouldn't we worry about that?

The empirical risk is just a proxy for a Bayesian risk that we are not able to compute unless we know the {\em true} distribution of the training data. We don't really mean to optimize the empirical risk.

Bayesian inference boils down to a problem of the form:
$$
r(\theta) = \int p(x) L(x, \theta) dx 
$$
where $p(x)$ is the distribution of some variable of interest and $\theta$ is the parameter to be learned. Because we don't know $p(x)$ in agnostic learning context, or we know it but it's intractable, we may resort to an empirical version of the risk:
$$
\hat{r}(\theta) = \frac{1}{n} L(x_i, \theta).
$$

You can use the Laplace approximation at the global optimum, it might work just as well as VS and probably faster. However, finding the global optimum is super challenging in high dimension: it usually exists (under mild conditions), but is hopeless to track using differential calculus based optimization, which generally converges to a saddle point. This is a fundamental limitation of numerical optimization techniques. Needless to say, Laplace is bound to work poorly at a saddle point since the Hessian is not definite positive, hence what you will get is a degenerate Gaussian fit.

Yet, according to the Bernstein-von Mises theorem, we expect Bayesian posteriors to be nearly Gaussian in the limit of large samples. How many data points do we need in practice, we don't know, but we should be able to tell how well this approximation holds in the problem at hand.

If it does hold, then the target distribution is peaked around a well-defined global optimum. Finding the global optimum is then akin to fitting a Gaussian distribution.

If it doesn't hold, then finding the global optimum makes little sense because we may have multiple local optima spread across the domain. But fitting a Gaussian may still make sense as it can at least tell us that the asymptotic regime is not reached, meaning that more training data is probably needed.

The fact that we get many saddle points in practice suggests that we may lack data, however this restriction is not fundamental if we have an underlying generative model to produce the training data.
